Python as a machine learning language: experiences from the scikit-learn project

### 2013-04-28 paper.tex
elaborated on object conversion (still commented out)
start section on implementation

For a machine learning library, one of the most important decisions we had to
make is how to represent data.  Rather than reinventing the wheel, we opted for
NumPy multidimensional arrays \Citep{vanderwalt2011} for dense data and SciPy
sparse matrices for sparse data.  While those may seem like a ``bare''
representation of data when compared to more object-oriented representations
(Weka is an example of this style of representation, \citealp{hall2009weka}),
it brings the prime advantage of allowing us to rely on NumPy and SciPy's
\textit{vector operations} which are often orders of magnitude faster
than the corresponding Python loops,
while at the same time keeping the code short and readable.
Conversion to NumPy and SciPy formats is usually easy and many scientific users
of Python will already be used to such formats, since they are pervasive in
other scientific Python packages.  For tasks where input is likely to consist
of text files or semi-structured objects, we provide ``vectorizers'' -- objects
that efficiently convert such data to the NumPy or SciPy formats.
% FIXME: elaborate what is meant below
%Finally, offering a custom datatype for samples would require a conversion
%\textit{in any case}.

@@

>>>>

For a machine learning library, one of the most important decisions we had to
make is how to represent data.  Rather than reinventing the wheel, we opted for
NumPy multidimensional arrays \Citep{vanderwalt2011} for dense data and SciPy
sparse matrices for sparse data.  While those may seem like a ``bare''
representation of data when compared to more object-oriented representations
(Weka is an example of this style of representation, \citealp{hall2009weka}),
it brings the prime advantage of allowing us to rely on NumPy and SciPy's
\textit{vector operations} which are often orders of magnitude faster
than the corresponding Python loops,
while at the same time keeping the code short and readable.
Conversion to NumPy and SciPy formats is usually easy and many scientific users
of Python will already be used to such formats, since they are pervasive in
other scientific Python packages.  For tasks where input is likely to consist
of text files or semi-structured objects, we provide ``vectorizers'' -- objects
that efficiently convert such data to the NumPy or SciPy formats.
% FIXME: elaborate what is meant below
% Lars: elaborated, but I'm not sure how important this is.
% It mostly reflects experiences with OO libraries that require conversions
% between types just to pass information to a method.
%Finally, offering a custom datatype for samples would require a conversion
%\textit{in any case}, since such a datatype would be toolkit-specific
%and no program would store its data using such a type
%if it weren't from the outset designed to use scikit-learn.

@@

\section{Implementation}

scikit-learn is primarily implemented in Python and Cython
\citep{behnel2011cython},
a language that extends Python with static typing
and a compiler that produces C extension modules
for the Python runtime system.
In addition, it includes (modified versions of)
the \textsf{LIBSVM} and \textsf{LIBLINEAR} libraries
used for training support vector machines
and logistic regression models \citep{chang2011libsvm, fan2008}.
These libraries, both written in C{}\verb!++!,
and wrapped using Cython modules.

As much as possible is implemented in ``pure'' Python,
to avoid the compile-edit-debug cycle needed for Cython
and to keep the codebase readable for as large an audience as possible.
However, since the speed of the current Python implementation
is not sufficient for numeric programming
and NumPy's vector operations are not appropriate for all use cases
(among other factors, because they often have linear space requirements),
some of the core algorithms have to be implemented in Cython.\footnote{
  The only Python implementation currently supported by scikit-learn
  is the reference implementation, CPython.
  The alternative Python implementation PyPy \citep{bolz2009tracing}
  promises to solve this problem by JIT compilation,
  but does not yet interoperate with NumPy and SciPy.
}
Examples include the aforementioned *k*-means implementation,
stochastic gradient descent for linear models
and some graph-based clustering algorithms.
